\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{psfig}

\begin{document}
\baselineskip=16pt
\textheight=9.6in
\parindent=0pt 
\def\sk {\hskip .5cm}
\def\skv {\vskip .12cm}
\def\cos {\mbox{cos}}
\def\sin {\mbox{sin}}
\def\tan {\mbox{tan}}
\def\intl{\int\limits}
\def\lm{\lim\limits}
\newcommand{\frc}{\displaystyle\frac}
\def\xbf{{\mathbf x}}
\def\fbf{{\mathbf f}}
\def\gbf{{\mathbf g}}

\def\dbA{{\mathbb A}}
\def\dbB{{\mathbb B}}
\def\dbC{{\mathbb C}}
\def\dbD{{\mathbb D}}
\def\dbE{{\mathbb E}}
\def\dbF{{\mathbb F}}
\def\dbG{{\mathbb G}}
\def\dbH{{\mathbb H}}
\def\dbI{{\mathbb I}}
\def\dbJ{{\mathbb J}}
\def\dbK{{\mathbb K}}
\def\dbL{{\mathbb L}}
\def\dbM{{\mathbb M}}
\def\dbN{{\mathbb N}}
\def\dbO{{\mathbb O}}
\def\dbP{{\mathbb P}}
\def\dbQ{{\mathbb Q}}
\def\dbR{{\mathbb R}}
\def\dbS{{\mathbb S}}
\def\dbT{{\mathbb T}}
\def\dbU{{\mathbb U}}
\def\dbV{{\mathbb V}}
\def\dbW{{\mathbb W}}
\def\dbX{{\mathbb X}}
\def\dbY{{\mathbb Y}}
\def\dbZ{{\mathbb Z}}

\def\la{{\langle}}
\def\ra{{\rangle}}

\def\Aut{{\rm Aut}}
\def\Inn{{\rm Inn}}

\bf\centerline{Homework \#4. }\rm
\vskip .1cm
{\bf Approximate plan for next week: } Direct and semi-direct products and further
applications of Sylow theorems (5.4, 5.5).

\vskip .1cm
\centerline{\bf Problems, to be submitted by Thursday, September 22nd}
\skv
{\bf 1.} Let $G$ be a finite group, $P$ a Sylow $p$-subgroup of $G$ for some $p$
and $H$ a subgroup of $G$ such that $N_{G}(P)\subseteq H\subseteq G$.
Prove that $N_H(P)=N_G(P)$ and $[G:H]\equiv 1\mod p$.
\skv

{\bf 2.} Prove that a group of order $200=8\cdot 25$ is not simple.
\skv
{\bf 3.} Prove that a group of order $132=3\cdot 4\cdot 11$ is not simple.
\skv

{\bf 4.} Let $G$ be a group of order $105=3\cdot 5\cdot 7$.
\begin{itemize}
\item[(a)] Prove that $G$ has a normal Sylow $5$-subgroup OR a normal Sylow $7$-subgroup.
\item[(b)] Use (a) to prove that $G$ has a normal subgroup of order $35$. Deduce that $G$ has
a normal Sylow $5$-subgroup AND a normal Sylow $7$-subgroup. 
\end{itemize} 
\skv
{\bf 5.} $\empty$
\begin{itemize}
\item[(a)] Let $G$ be an abelian group. Prove that the set of elements in $G$ which
have finite order is a subgroup of $G$. This subgroup is called the
{\it torsion subgroup } of $G$.

\item[(b)] Give an example of a group where the set of elements of finite order
is not a subgroup. {\bf Note:} Obviously, the group has to be infinite
and non-abelian by (a). 
\end{itemize}
\end{document}
