\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{psfig}

\begin{document}
\baselineskip=16pt
\textheight=9in
\textwidth=6in
\parindent=0pt
\def\sk {\hskip .5cm}
\def\skv {\vskip .12cm}
\def\cos {\mbox{cos}}
\def\sin {\mbox{sin}}
\def\tan {\mbox{tan}}
\def\intl{\int\limits}
\def\lm{\lim\limits}
\newcommand{\frc}{\displaystyle\frac}
\def\xbf{{\mathbf x}}
\def\fbf{{\mathbf f}}
\def\gbf{{\mathbf g}}

\def\dbA{{\mathbb A}}
\def\dbB{{\mathbb B}}
\def\dbC{{\mathbb C}}
\def\dbD{{\mathbb D}}
\def\dbE{{\mathbb E}}
\def\dbF{{\mathbb F}}
\def\dbG{{\mathbb G}}
\def\dbH{{\mathbb H}}
\def\dbI{{\mathbb I}}
\def\dbJ{{\mathbb J}}
\def\dbK{{\mathbb K}}
\def\dbL{{\mathbb L}}
\def\dbM{{\mathbb M}}
\def\dbN{{\mathbb N}}
\def\dbO{{\mathbb O}}
\def\dbP{{\mathbb P}}
\def\dbQ{{\mathbb Q}}
\def\dbR{{\mathbb R}}
\def\dbS{{\mathbb S}}
\def\dbT{{\mathbb T}}
\def\dbU{{\mathbb U}}
\def\dbV{{\mathbb V}}
\def\dbW{{\mathbb W}}
\def\dbX{{\mathbb X}}
\def\dbY{{\mathbb Y}}
\def\dbZ{{\mathbb Z}}

\def\la{{\langle}}
\def\ra{{\rangle}}

\def\Aut{{\rm Aut}}
\def\Inn{{\rm Inn}}

\bf\centerline{Homework \#7. }\rm
\vskip .1cm
{\bf Plan for next week:} Free groups (continued) and presentations of groups by generators and relators (\S~6.3), start ring theory (\S~7.1).
\vskip .1cm
\centerline{\bf Problems, to be submitted by Thursday, October, 20th}
\skv
{\bf 1.} Let $n\geq 3$ and $G=D_{2n}$, the dihedral group of order $2n$.

(a) Prove that $G$ contains a subgroup isomorphic to $D_{2k}$
for any $k\mid n$.

(b) Prove that the dihedral group $G$ is nilpotent if and only
if $n$ is power of $2$.
\skv
{\bf 2.} (a) Let $R$ be an associative ring with $1$, and let $a,b\in R$ be such that
$1+a$ and $1+b$ are invertible. Prove the following formula 
$$(1+a)^{-1}(1+b)^{-1}(1+a)(1+b)=1+(1+a)^{-1}(1+b)^{-1}(ab-ba).$$

(b) Let $R$ be an associative ring with $1$ and $n\geq 2$ be an integer, and let
$U_n(R)$ be the upper unitriangular subgroup of $GL_n(R)$. Prove that $U_n(R)$
is niplotent of class $n-1$ (we briefly outlined the proof in class). Note that you
will need to apply (a) not to $R$ itself but to the ring of $n\times n$ matrices over $R$.
\skv
{\bf 3.} (a) Let $n\in\dbN$ be an integer, and suppose that for every non-prime divisor $m$ of $n$ there are no simple groups of order $m$. Prove that any group of order $n$ is solvable.
\skv
(b) Prove that any group of order $p^k q$, where $p>q$ are distinct primes, is solvable.

\skv
{\bf 4.} Problems 31 and 32 on page 200 of DF. {\bf Note:} Problem 31 follows
very easily from Lemma~11.2 (about the structure of minimal normal subgroups).
\skv
{\bf 5.} (a) Let $A$ and $B$ be finitely generated groups. Prove that the wreath product $A wr B$
is also finitely generated.
{\bf Hint:} Recall that $A wr B=C\rtimes B$ where $C=\oplus_{b\in B}A_b$ (with each $A_b\cong A$). Let $S$
be a generating set for $A$, $T$ a generating set for $B$, fix $b\in B$, and let $S_b$ be the image
of $S$ under an isomorphism $A\to A_b$. Prove that $S_b\cup T$ generates $A wr B$.
\skv
(b) Use (a) to give a simple example showing that a subgroup of a finitely generated group
may not be finitely generated.
\skv
{\bf 6. }(a) (practice) In class we outlined the proof of the fact that $\dbZ_p wr \dbZ_p$
is isomorphic to the Sylow $p$-subgroup of $S_{p^2}$. Fill in the details of that proof.
\skv
(b) (bonus) Realize the Sylow $p$-subgroup of $S_{p^3}$ in terms of wreath products and prove
your answer (you may skip some technical details). 
\end{document}
