\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{psfig}

\begin{document}
\baselineskip=15pt
\textheight=9in
\parindent=0pt
\def\sk {\hskip .5cm}
\def\skv {\vskip .07cm}
\def\cos {\mbox{cos}}
\def\sin {\mbox{sin}}
\def\tan {\mbox{tan}}
\def\intl{\int\limits}
\def\lm{\lim\limits}
\newcommand{\frc}{\displaystyle\frac}
\def\xbf{{\mathbf x}}
\def\fbf{{\mathbf f}}
\def\gbf{{\mathbf g}}

\def\Ker{{\rm Ker\,}}
\def\Gal{{\rm Gal\,}}
\def\phi{\varphi}

\def\dbA{{\mathbb A}}
\def\dbB{{\mathbb B}}
\def\dbC{{\mathbb C}}
\def\dbD{{\mathbb D}}
\def\dbE{{\mathbb E}}
\def\dbF{{\mathbb F}}
\def\dbG{{\mathbb G}}
\def\dbH{{\mathbb H}}
\def\dbI{{\mathbb I}}
\def\dbJ{{\mathbb J}}
\def\dbK{{\mathbb K}}
\def\dbL{{\mathbb L}}
\def\dbM{{\mathbb M}}
\def\dbN{{\mathbb N}}
\def\dbO{{\mathbb O}}
\def\dbP{{\mathbb P}}
\def\dbQ{{\mathbb Q}}
\def\dbR{{\mathbb R}}
\def\dbS{{\mathbb S}}
\def\dbT{{\mathbb T}}
\def\dbU{{\mathbb U}}
\def\dbV{{\mathbb V}}
\def\dbW{{\mathbb W}}
\def\dbX{{\mathbb X}}
\def\dbY{{\mathbb Y}}
\def\dbZ{{\mathbb Z}}

\def\Aut{{\rm Aut}}
\def\deg{{\rm deg}}

\def\la{{\langle}}
\def\ra{{\rangle}}

\bf\centerline{Homework Assignment \# 8. }\rm

{\bf Plan for the next week:} Galois extensions and Galois groups (online lecture 19, Section 14.1 in DF), Galois correspondence (online lectures 20 and 21, Section 14.2 in DF).
\skv
\bf\centerline{Problems, to be submitted by Thu, March 28th. }\rm
\skv
{\bf Problem 1:} \rm Let $p$ be a prime, $K=\mathbb F_p(s,t)$, the field of rational functions over $\dbF_p$ in two variables, and
let $F=\mathbb F_p(s^p,t^p)$. Prove that the extension $K/F$ cannot be generated by a single element.
\skv
{\bf Problem 2:} \rm
Prove the interesting part of Corollary 18.7 form online notes: if $K/L/F$ is a tower of algebraic extensions and $K/L$ and $L/F$ are separable, then $K/F$ is separable (see online notes for a hint). Give a detailed argument.
\skv
{\bf Problem 3:} \rm Let $p_1,\ldots, p_n$ be distinct primes. Let $\alpha_i=\sqrt[p_i]{p_i}$ and $\alpha=\sum\limits_{i=1}^n \alpha_i$. Prove that $\dbQ(\alpha_1,\ldots,\alpha_n)=\dbQ(\alpha)$. {\bf Hint:} this is somewhat similar to the example we discussed at the end of Lecture~18, but certain details of the proof are different.
\skv
{\bf Problem 4:} \rm  Let $p$ be a prime, $n\geq 2$ an integer, $f(x)=x^n-p$, and let
$K\subset \dbC$ be the splitting field for $f(x)$ over $\dbQ$.
Recall that $K=\dbQ(\sqrt[n]{p},\omega_n)$ where
$\omega_n=e^{2\pi i/n}$.
\begin{itemize}
\item[(a)] Describe the elements of $\Gal(K/\dbQ)$ by their actions on $\sqrt[n]{p}$
and $\omega_n$.
\item[(b)] Let $G=\dbZ/n\dbZ\rtimes (\dbZ/n\dbZ)^{\times}$ where 
$(\dbZ/n\dbZ)^{\times}$
acts on $\dbZ/n\dbZ$ by multiplication. Prove that $\Gal(K/\dbQ)$ is isomorphic
to a subgroup of $G$ and describe an explicit embedding. 
\item[(c)] (optional) Under which conditions on $n$ and $p$ is this embedding an isomorphism? 
You are not asked for necessary and sufficient conditions; just the most general sufficient
conditions that you can justify.
\end{itemize}

{\bf Problem 5:} \rm Let $F$ be a field, let $f(x)\in F[x]$
be an irreducible (over $F$) separable polynomial of degree $n$,
and let $K$ be a splitting field of $f(x)$.

\begin{itemize}
\item[(a)] Prove that $|\Gal(K/F)|$ is a multiple of $n$ and divides $n!$

\item[(b)] Let $n=3$. Prove that $\Gal(K/F)$ is isomorphic to either $\dbZ/3\dbZ$ or $S_3$.

\item[(c)] Let $n=4$ and assume that $|\Gal(K/F)|=8$.
Determine the isomorphism class of $\Gal(K/F)$.
\end{itemize}
\end{document} 