\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{psfig}

\begin{document}
\baselineskip=15pt
\textheight=9in
\parindent=0pt
\def\sk {\hskip .5cm}
\def\skv {\vskip .07cm}
\def\cos {\mbox{cos}}
\def\sin {\mbox{sin}}
\def\tan {\mbox{tan}}
\def\intl{\int\limits}
\def\lm{\lim\limits}
\newcommand{\frc}{\displaystyle\frac}
\def\xbf{{\mathbf x}}
\def\fbf{{\mathbf f}}
\def\gbf{{\mathbf g}}

\def\Ker{{\rm Ker\,}}
\def\Gal{{\rm Gal\,}}
\def\phi{\varphi}

\def\dbA{{\mathbb A}}
\def\dbB{{\mathbb B}}
\def\dbC{{\mathbb C}}
\def\dbD{{\mathbb D}}
\def\dbE{{\mathbb E}}
\def\dbF{{\mathbb F}}
\def\dbG{{\mathbb G}}
\def\dbH{{\mathbb H}}
\def\dbI{{\mathbb I}}
\def\dbJ{{\mathbb J}}
\def\dbK{{\mathbb K}}
\def\dbL{{\mathbb L}}
\def\dbM{{\mathbb M}}
\def\dbN{{\mathbb N}}
\def\dbO{{\mathbb O}}
\def\dbP{{\mathbb P}}
\def\dbQ{{\mathbb Q}}
\def\dbR{{\mathbb R}}
\def\dbS{{\mathbb S}}
\def\dbT{{\mathbb T}}
\def\dbU{{\mathbb U}}
\def\dbV{{\mathbb V}}
\def\dbW{{\mathbb W}}
\def\dbX{{\mathbb X}}
\def\dbY{{\mathbb Y}}
\def\dbZ{{\mathbb Z}}

\def\Aut{{\rm Aut}}
\def\deg{{\rm deg}}

\def\la{{\langle}}
\def\ra{{\rangle}}

\bf\centerline{Homework Assignment \# 10. }\rm

{\bf Plan for the next week:} Section 10.5 (exact sequences and injective and projective modules)
\skv
\skv
\bf\centerline{Problems, to be submitted by Thu, April 18th. }\rm
\skv
\skv
{\bf Problem 1:} Let $f(x)$ and $g(x)$ be irreducible polynomials in $\mathbb F_p[x]$
of the same degree and let $F=\mathbb F_p[x]/(f(x))$. Prove that $g(x)$ splits completely over $F$.
\skv
{\bf Problem 2:} Let $p$ be a prime, $n$ a positive integer and
$\Phi_n(x)=x^{p^n}-x\in \mathbb F_p[x]$. Prove that $\Phi_n(x)$ is equal to the product
of all monic irreducible polynomials in $\mathbb F_p[x]$ whose degrees divide $n$
(where each polynomial occurs with multiplicity one).
\skv
\skv
{\bf Problem 3:} Prove the following analogue of Kummer's theorem for abelian extensions:
Let $n\in\dbN$ and let $F$ be a field containing primitive $n^{\rm th}$ root of unity.
\begin{itemize}
\item[(a)] Let $K/F$ be a finite Galois extension such that $\Gal(K/F)$ is abelian
of exponent $n$. Then there exists $a_1,\ldots, a_t\in K$ s.t.
$K=F(\sqrt[n]{a_1},\ldots,\sqrt[n]{a_t})$, or more precisely, there exists
$\alpha_1,\ldots, \alpha_t\in K$ s.t.
$K=F(\alpha_1,\ldots,\alpha_t)$ and $\alpha_i^n\in F$ for all $i$.

\item[(b)] Conversely, suppose that $K=F(\sqrt[n]{a_1},\ldots,\sqrt[n]{a_t})$
for some $a_1,\ldots, a_t\in F$. Prove that $K/F$ is Galois, and $\Gal(K/F)$
is abelian of exponent $n$.
\end{itemize}
\skv
{\bf Problem 4:} (practice). DF, problem 18 on p.583. {\bf Note:}  Make sure to solve this problem if you did not solve  
Problem~17 completely correctly.

\skv
{\bf Problem 5:} DF, problem 21 on p.583.
\skv
{\bf Problem 6:} Let $p$ be a prime, $K=\mathbb F_p(t)$, the field of rational functions in one variable over $\dbF_p$.
Let $\sigma:K\to K$ be the unique automorphism of $K$ such that $\sigma(t)=t+1$ and $G=\langle \sigma\rangle$ (clearly
$G$ is cyclic of order $p$). Let $F=K^G$. Find an explicit element $s$ such that $F=\dbF_p(s)$ and prove your answer.
\skv
{\bf Problem 7:} DF, problem 9 on p.636.
\skv


\end{document} 