\documentclass[12pt]{amsart}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{url}
\usepackage{hyperref}
%\usepackage{psfig}

\begin{document}
\baselineskip=15pt
\textheight=8.4in
\parindent=0pt
\def\sk {\hskip .5cm}
\def\skv {\vskip .12cm}
\def\cos {\mbox{cos}}
\def\sin {\mbox{sin}}
\def\tan {\mbox{tan}}
\def\intl{\int\limits}
\def\lm{\lim\limits}
\newcommand{\frc}{\displaystyle\frac}
\def\xbf{{\mathbf x}}
\def\fbf{{\mathbf f}}
\def\gbf{{\mathbf g}}

\def\Ker{{\rm Ker\,}}
\def\phi{\varphi}

\def\dbA{{\mathbb A}}
\def\dbB{{\mathbb B}}
\def\dbC{{\mathbb C}}
\def\dbD{{\mathbb D}}
\def\dbE{{\mathbb E}}
\def\dbF{{\mathbb F}}
\def\dbG{{\mathbb G}}
\def\dbH{{\mathbb H}}
\def\dbI{{\mathbb I}}
\def\dbJ{{\mathbb J}}
\def\dbK{{\mathbb K}}
\def\dbL{{\mathbb L}}
\def\dbM{{\mathbb M}}
\def\dbN{{\mathbb N}}
\def\dbO{{\mathbb O}}
\def\dbP{{\mathbb P}}
\def\dbQ{{\mathbb Q}}
\def\dbR{{\mathbb R}}
\def\dbS{{\mathbb S}}
\def\dbT{{\mathbb T}}
\def\dbU{{\mathbb U}}
\def\dbV{{\mathbb V}}
\def\dbW{{\mathbb W}}
\def\dbX{{\mathbb X}}
\def\dbY{{\mathbb Y}}
\def\dbZ{{\mathbb Z}}

\def\Aut{{\rm Aut}}
\def\Gal{{\rm Gal}}
\def\exp{{\rm exp}}
\def\det{{\rm det}}
\def\tr{{\rm tr}}

\def\la{{\langle}}
\def\ra{{\rangle}}
\def\rk{{\rm rk}}


\bf\centerline{Homework Assignment \# 10}\rm
\vskip .2cm
{\bf Plan for the class on Apr 13:} We will finish the discussion of cyclic extensions (end of online lecture 23) and then talk about solvable
and radical extensions (online lecture 24, section 14.7 in DF).
\vskip .1cm
Here and in all future assignments ``online'' or ``online notes'' refers to Algebra-II lectures posted on my Spring 2010 Algebra-II webpage
\vskip .1cm
\centerline{\url{http://people.virginia.edu/~mve2x/7752_Spring2010/}}
\vskip .1cm

%{\bf Note on hints:} All hints are given at the end of the %assignment, each on a separate page.
%Problems (or parts of problems) for which hint is available are %marked with *.

\vskip .3cm

\bf\centerline{Problems, due by 11:59pm on Friday, April 16th.}\rm
\vskip .1cm
{\bf Problem 0:} (practice). DF, problem 18 on p.583. {\bf Note:}  Make sure to solve this problem if you did not solve  
Problem~17 correctly.
\skv
{\bf Problem 1:} DF, problem 21 on p.583.

\skv
{\bf Problem 2:} Prove the additive form of Hilbert's Theorem 90 (DF, problem 26 on p.584). The proof is very similar to the proof
of the multiplicative form we established in class. 
\skv
{\bf Problem 3:} DF, problems 4 and 5 on p.636.  
\skv
{\bf Problem 4:} Let $p$ be a prime, $K=\mathbb F_p(t)$, the field of rational functions in one variable over $\dbF_p$.
Let $\sigma:K\to K$ be the unique automorphism of $K$ such that $\sigma(t)=t+1$ and $G=\langle \sigma\rangle$ (clearly
$G$ is cyclic of order $p$). Let $F=K^G$. Find an explicit element $s$ such that $F=\dbF_p(s)$ and prove your answer.
\skv
{\bf Problem 5:} 
Let $E/F$ be a field extension, let $K_1$ and $K_2$ be subfields of $E$ containing $F$,
and assume that the extensions $K_1/F$ and $K_2/F$ are finite. Let $K_1 K_2$ be the composite
field of $K_1$ and $K_2$. Prove that the $F$-algebra $K_1\otimes_F K_2$ is a field if and only if
$$[K_1 K_2: F]=[K_1: F][K_2: F].$$ {\bf Hint:} Use the universal property of tensor products to relate
$K_1\otimes_F K_2$ and $K_1 K_2$.
\end{document}
