\documentclass[12pt]{amsart}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{url}
\usepackage{hyperref}
%\usepackage{psfig}

\begin{document}
\baselineskip=15pt
\textheight=8.4in
\parindent=0pt
\def\sk {\hskip .5cm}
\def\skv {\vskip .12cm}
\def\cos {\mbox{cos}}
\def\sin {\mbox{sin}}
\def\tan {\mbox{tan}}
\def\intl{\int\limits}
\def\lm{\lim\limits}
\newcommand{\frc}{\displaystyle\frac}
\def\xbf{{\mathbf x}}
\def\fbf{{\mathbf f}}
\def\gbf{{\mathbf g}}

\def\Ker{{\rm Ker\,}}
\def\phi{\varphi}

\def\dbA{{\mathbb A}}
\def\dbB{{\mathbb B}}
\def\dbC{{\mathbb C}}
\def\dbD{{\mathbb D}}
\def\dbE{{\mathbb E}}
\def\dbF{{\mathbb F}}
\def\dbG{{\mathbb G}}
\def\dbH{{\mathbb H}}
\def\dbI{{\mathbb I}}
\def\dbJ{{\mathbb J}}
\def\dbK{{\mathbb K}}
\def\dbL{{\mathbb L}}
\def\dbM{{\mathbb M}}
\def\dbN{{\mathbb N}}
\def\dbO{{\mathbb O}}
\def\dbP{{\mathbb P}}
\def\dbQ{{\mathbb Q}}
\def\dbR{{\mathbb R}}
\def\dbS{{\mathbb S}}
\def\dbT{{\mathbb T}}
\def\dbU{{\mathbb U}}
\def\dbV{{\mathbb V}}
\def\dbW{{\mathbb W}}
\def\dbX{{\mathbb X}}
\def\dbY{{\mathbb Y}}
\def\dbZ{{\mathbb Z}}

\def\Aut{{\rm Aut}}
\def\Gal{{\rm Gal}}
\def\exp{{\rm exp}}
\def\det{{\rm det}}
\def\tr{{\rm tr}}

\def\la{{\langle}}
\def\ra{{\rangle}}
\def\rk{{\rm rk}}


\bf\centerline{Homework Assignment \# 8}\rm
\vskip .2cm
{\bf Plan for the week of Mar 29:} The main topic will be Galois correspondence (online lectures 20 and 21, Section 14.2 in DF). At the end of Thu Lecture we will either talk about cyclotomic fields (13.6) or continue
talking about finite fields, picking up where we left off in Lecture 14 on Tue, March 23 (online lecture 22,
Section 14.3 in DF)
\vskip .1cm
Here and in all future assignments ``online'' or ``online notes'' refers to Algebra-II lectures posted on my Spring 2010 Algebra-II webpage
\vskip .1cm
\centerline{\url{http://people.virginia.edu/~mve2x/7752_Spring2010/}}
\vskip .1cm

%{\bf Note on hints:} All hints are given at the end of the %assignment, each on a separate page.
%Problems (or parts of problems) for which hint is available are %marked with *.

\vskip .3cm

\bf\centerline{Problems, due by 11:59pm on Friday, April 2nd.}\rm
\vskip .1cm
\skv
{\bf Problem 1:} \rm Let $p$ be a prime, $K=\mathbb F_p(s,t)$, the field of rational functions over $\dbF_p$ in two variables, and
let $F=\mathbb F_p(s^p,t^p)$. Prove that the extension $K/F$ cannot be generated by a single element.

\skv
{\bf Problem 2:} \rm Let $p_1,\ldots, p_n$ be distinct primes. Let $\alpha_i=\sqrt[p_i]{p_i}$ and $\alpha=\sum\limits_{i=1}^n \alpha_i$. Prove that $\dbQ(\alpha_1,\ldots,\alpha_n)=\dbQ(\alpha)$. {\bf Hint:} this is somewhat similar to the example we discussed at the end of Lecture~18, but certain details of the proof are different.
\skv
{\bf Problem 3:} \rm  Let $p$ be a prime, $n\geq 2$ an integer, $f(x)=x^n-p$, and let
$K\subset \dbC$ be the splitting field for $f(x)$ over $\dbQ$.
Recall that $K=\dbQ(\sqrt[n]{p},\omega_n)$ where
$\omega_n=e^{2\pi i/n}$.
\begin{itemize}
\item[(a)] Describe the elements of $\Gal(K/\dbQ)$ by their actions on $\sqrt[n]{p}$
and $\omega_n$.
\item[(b)] Let $G=\dbZ/n\dbZ\rtimes (\dbZ/n\dbZ)^{\times}$ where 
$(\dbZ/n\dbZ)^{\times}$
acts on $\dbZ/n\dbZ$ by multiplication. Prove that $\Gal(K/\dbQ)$ is isomorphic
to a subgroup of $G$ and describe an explicit embedding. 
\item[(c)] (optional) Under which conditions on $n$ and $p$ is this embedding an isomorphism? 
You are not asked for necessary and sufficient conditions; just the most general sufficient
conditions that you can justify.
\end{itemize}
\skv
{\bf Problem 4:} \rm Let $F$ be a field, let $f(x)\in F[x]$
be an irreducible (over $F$) separable polynomial of degree $n$,
and let $K$ be a splitting field of $f(x)$.

\begin{itemize}
\item[(a)] Prove that $|\Gal(K/F)|$ is a multiple of $n$ and divides $n!$

\item[(b)] Let $n=3$. Prove that $\Gal(K/F)$ is isomorphic to either $\dbZ/3\dbZ$ or $S_3$.

\item[(c)] Assume that $F=\dbR$ (real numbers) and $f(x)$ has a non-real root. Prove that
$\Gal(K/F)$ is isomorphic to $S_3$.
\end{itemize}
\skv
{\bf Problem 5:} Let $f(x)$ and $g(x)$ be irreducible polynomials in $\mathbb F_p[x]$
of the same degree and let $F=\mathbb F_p[x]/(f(x))$. Prove that $g(x)$ splits completely over $F$.
\skv
{\bf Problem 6:} Let $p$ be a prime, $n$ a positive integer and
$\Phi_n(x)=x^{p^n}-x\in \mathbb F_p[x]$. Prove that $\Phi_n(x)$ is equal to the product
of all monic irreducible polynomials in $\mathbb F_p[x]$ whose degrees divide $n$
(where each polynomial occurs with multiplicity one).
\skv
\end{document}
