\documentclass[12pt]{amsart}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{psfig}

\newtheorem* {Theorem}    {Theorem}
\newtheorem* {Lemma}    {Lemma}


\begin{document}
 \pagenumbering{gobble}
\baselineskip=16pt
\textheight=8.5in
\textwidth=6.5in
%\parindent=0pt 
\def\sk {\hskip .5cm}
\def\skv {\vskip .08cm}
\def\cos {\mbox{cos}}
\def\sin {\mbox{sin}}
\def\tan {\mbox{tan}}
\def\intl{\int\limits}
\def\lm{\lim\limits}
\newcommand{\frc}{\displaystyle\frac}
\def\xbf{{\mathbf x}}
\def\fbf{{\mathbf f}}
\def\gbf{{\mathbf g}}

\def\dbA{{\mathbb A}}
\def\dbB{{\mathbb B}}
\def\dbC{{\mathbb C}}
\def\dbD{{\mathbb D}}
\def\dbE{{\mathbb E}}
\def\dbF{{\mathbb F}}
\def\dbG{{\mathbb G}}
\def\dbH{{\mathbb H}}
\def\dbI{{\mathbb I}}
\def\dbJ{{\mathbb J}}
\def\dbK{{\mathbb K}}
\def\dbL{{\mathbb L}}
\def\dbM{{\mathbb M}}
\def\dbN{{\mathbb N}}
\def\dbO{{\mathbb O}}
\def\dbP{{\mathbb P}}
\def\dbQ{{\mathbb Q}}
\def\dbR{{\mathbb R}}
\def\dbS{{\mathbb S}}
\def\dbT{{\mathbb T}}
\def\dbU{{\mathbb U}}
\def\dbV{{\mathbb V}}
\def\dbW{{\mathbb W}}
\def\dbX{{\mathbb X}}
\def\dbY{{\mathbb Y}}
\def\dbZ{{\mathbb Z}}

\def\la{{\langle}}
\def\ra{{\rangle}}
\def\Ker{{\rm Ker}}
\def\Aut{{\rm Aut}}
\def\rk{{\rm rk}}
\def\summ{{\sum\limits}}
\def\phi{{\varphi}}

\bf\centerline{Math 8851. Homework \#2. To be completed by Thu, Feb 9}\rm
\vskip .1cm

1. Let $G$ be a finitely presented group. Define the deficiency
of $G$, denoted by $def(G)$, to be the supremum of the quantity
$|X|-|R|$ where $(X,R)$ runs over all presentations of $G$
by generators and relators.

\begin{itemize}
\item[(a)] Let $H$ be a subgroup of $G$ of index $n$.
Prove that $def(H)\geq 1+(def(G)-1)\cdot n$.

\item[(b)] Prove that $def(G)\leq d(G^{ab})$ where
$G^{ab}=G/[G,G]$ is the abelianization of $G$ and
$d(\cdot)$ denoted the minimal number of generators. 
In particular, this implies that $def(G)<\infty$.

{\bf Hint:} Suppose that $G=\la X | R\ra$ where
$X=\{x_1,\ldots, x_n\}$ and $R=\{r_1,\ldots, r_m\}$. Show that $G^{ab}$
is isomorphic to $\dbZ^n/M$ where $M$ is a subgroup of $\dbZ^n$
generated by $m$ elements.

\end{itemize}

\vskip .1cm

2. The goal of this problem is to fill in the details of the proof of
Hall's theorem (extensions of finitely presented groups are finitely presented).
So, suppose we are given a short exact sequence of groups
$$1\to N \to G\stackrel{\pi}{\rightarrow} Q\to 1$$
where $N$ and $Q$ are finitely presented. Our goal is to prove that
$G$ is finitely presented.

As in class, we can assume that $Q=G/N$ and $\pi(g)=gN$ for all $g\in G$.
Start by choosing finite presentations $N=\la Y | A\ra$ and 
$Q=\la Z | B\ra$ where $Y=\{y_1,\ldots, y_k\}$ and $Z=\{z_1,\ldots, z_m\}$.
For each $1\leq i\leq m$ choose $\widetilde z_i\in G$ such that 
$\pi(\widetilde z_i)=z_i$, and let 
$\widetilde Z=\{\widetilde z_1,\ldots, \widetilde z_m\}$. We claim
that $Y\cup \widetilde Z$ is a generating set of $G$.

Now take any $b\in B$ and think of $b$ as a word in $z_1,\ldots, z_m$.
Let $\widetilde b= b(\widetilde z_1,\ldots, \widetilde z_m)$ be the word
obtained from $b$ by replacing each $z_i$ by $\widetilde z_i$, and let
$\widetilde b_{\widetilde G}$ be the element of $\widetilde G$ represented by $\widetilde b$.
Then $\pi(\widetilde b_{\widetilde G})=b(z_1,\ldots, z_n)=1$ in $G$, so $\widetilde b_{\widetilde G}\in N$.
Since $N$ is generated by $y_1,\ldots, y_k$, there exists a word $w_b$ in $y_1,\ldots,
y_k$, such that the word $\widetilde b\cdot  w_b(y_1,\ldots, y_k)^{-1}$ is a relator of $G$
(with respect to the generating set $Y\cup \widetilde Z$). Let
$$\widetilde B=\{\widetilde b\cdot  w_b(y_1,\ldots, y_k)^{-1}: b\in B\}.$$

Next for each $1\leq i\leq m$ and $1\leq j\leq k$ the element
$\widetilde z_i y_j \widetilde z_i^{\,-1}$ lies in $N$ (since $N$ is normal in $G$),
so $\widetilde z_i y_j \widetilde z_i^{\,-1}=w_{ij}(y_1,\ldots, y_k)$ for some word $w_{ij}$ in $y_1,\ldots, y_k$. Let
$$C=
\{\widetilde z_i y_j \widetilde z_i^{\,-1}\cdot w_{ij}(y_1,\ldots, y_k)^{-1}: 1\leq i\leq m, 1\leq j\leq k\},$$
and let $\widetilde G$ be the group given by the presentation
$\la Y\cup \widetilde Z | R\ra$ where $R= A\cup \widetilde B\cup C$.
By construction each element of $R$ is a relator of $G$, so by von Dyck's theorem there exists a homomorphism $\phi: \widetilde G\to G$ which sends
every $x\in Y\cup \widetilde Z$ considered as an element of $\widetilde G$
to the same $x$ considered as an element of $G$. Our goal is to show that
$\phi$ is an isomorphism.

Now the actual problem starts.

\begin{itemize}
\item[(a)] Prove that $Y\cup \widetilde Z$ is indeed a generating set for $G$. Deduce that $\phi$ is surjective.
\item[(b)] Now let $\widetilde N$ be the subgroup of $\widetilde G$ generated
by $Y$. Show that $\widetilde N$ is normal in $\widetilde G$ and 
$\phi$ maps $\widetilde N$ isomorphically onto $N$ (to prove that
$\phi$ restricted to $\widetilde N$ is an isomoprphism use von Dyck's theorem
to construct a natural map $\psi: N\to \widetilde N$ and show that it is inverse to $\phi$ restricted to $N$).
\item[(c)] Deduce from (b) that $\phi$ induces a natural map
$\overline \phi: \widetilde G/ \widetilde N \to G/N$ and then prove that
$\overline \phi$ is an isomorphism similarly to (b).
\item[(d)] Finally deduce from (b) and (c) that $\phi:\widetilde G\to G$
is an isomorphism.
\end{itemize}
\skv
3. Consider the following condition (*) on a group $G$:
\begin{itemize}
\item[(*)] There exists a central extension $1\to N\to \widehat G\to G\to 1$ (that is, this sequence is exact and 
$N$ lies in $Z(\widehat G)$, the center of $\widehat G$) such that $\widehat G$ is finitely generated and $N$ is not finitely generated.
\end{itemize}
In Lecture~5 we proved that any group satisfying (*) is not finitely presented (note that such group is obviously finitely generated being a quotient
of the finitely generated group $\widehat G$).

\begin{itemize}
\item[(a)] Prove that $G$ satisfies (*) if and only if $G\cong F/N$ for some finitely generated group $F$ and a normal subgroup $N$
such that the quotient $N/[N,F]$ is infinitely generated (if $A$ and $B$ are subgroups of the same group, $[A,B]$ is the subgroup
generated by all commutators $[a,b]$ with $a\in A$, $b\in B$).
\item[(b)] Let $G$ be a finitely generated group, so that $G\cong F/N$ for some finitely generated group $F$ and a normal subgroup $N$.
For simplicity of notation we will assume below that $G=F/N$.

\skv
\skv

Prove that each of the following conditions implies the next one:
\begin{itemize}
\item[(i)] $G$ is finitely presented
\item[(ii)] $N^{ab}=N/[N,N]$ is finitely generated as a $G$-module where $G=F/N$ acts on $N^{ab}$ by conjugation:
$$(gN).(y[N,N])=gyg^{-1}[N,N]\mbox{ for all }g\in G, y\in N.$$
\item[(iii)] $N/[N,F]$ is finitely generated as a group.
\end{itemize}
\end{itemize}
Using the Hopf formula mentioned in class, it is not hard to show that (iii) holds if and only if $H_2(G,\dbZ)$ is finitely generated.
Once can also show that (ii) holds if and only if $G$ has property $(FP_2)$.
\skv
{\bf 4. } For any groups $A$ and $B$ their {\it wreath product} $A\, wr\, B$ is defined as follows.
Let $C=(\oplus_{a\in A}B)$, and think of elements of $C$ as functions $f:A\to B$ with finite support (that is, $f(a)=1_B$ for all but finitely many $a\in A$).
Define the homomorphism $\phi:A\to \Aut(C)$ by $$(\phi(x)f)(a)=f(x^{-1}a) \mbox{ for all } x,a\in A \mbox{ and }f\in C.$$ 
The wreath product $A\, wr\, B$ is the semidirect product $A\rtimes_{\phi} C$.

As usual, we can think of $A$ and $C$ as subgroups of $A\, wr\, B$. For each $a\in A$ and $b\in B$ let $b_a$ be the element of $C$
defined by $b_a(a)=b$ and $b_a(x)=1_B$ for all $x\in B\setminus \{a\}$.

\begin{itemize}
\item[(a)] Verify that $x b_a x^{-1}=b_{xa}$ for all $x,a\in A$ and $b\in B$.
\item[(b)] Let $S$ be a generating set for $A$ and $T$ a generating set for $B$.
Given $a\in A$, let $B_a=\{b_a: b\in B\}$ -- this is a subgroup of $C$ isomorphic to $B$; you should think of it as a copy of $B$
indexed by $a$ -- and let $T_a\subseteq B_a$ be the corresponding copy of $T$. Prove that
if we fix $x\in A$, then $T_x\cup S$ is a generating set for $A\, wr\, B$. Deduce
that the wreath product of finitely generated groups is finitely generated.
{\bf Hint:} First show that $\cup_{a\in A} T_a\cup S$ is a generating set for $A\, wr\, B$.
\item[(c)] Given a prime $p$, the group $G_p=\dbZ\, wr\, \dbZ/p\dbZ$ is called the $p$-lamplighter group. Prove that
$G_p$ has the following presentation by generators and relations:
$$\la x,y \mid y^p=1, [y,y^x]=1, [y, y^{x^2}]=1, [y, y^{x^3}]=1,\ldots\ra.$$
\end{itemize}
Here $g^h=h^{-1}gh$.
\end{document}



