\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{psfig}

\begin{document}
\baselineskip=15pt
\textheight=9in
\parindent=0pt
\def\sk {\hskip .5cm}
\def\skv {\vskip .07cm}
\def\cos {\mbox{cos}}
\def\sin {\mbox{sin}}
\def\tan {\mbox{tan}}
\def\intl{\int\limits}
\def\lm{\lim\limits}
\newcommand{\frc}{\displaystyle\frac}
\def\xbf{{\mathbf x}}
\def\fbf{{\mathbf f}}
\def\gbf{{\mathbf g}}

\def\Ker{{\rm Ker\,}}
\def\Gal{{\rm Gal\,}}
\def\phi{\varphi}

\def\dbA{{\mathbb A}}
\def\dbB{{\mathbb B}}
\def\dbC{{\mathbb C}}
\def\dbD{{\mathbb D}}
\def\dbE{{\mathbb E}}
\def\dbF{{\mathbb F}}
\def\dbG{{\mathbb G}}
\def\dbH{{\mathbb H}}
\def\dbI{{\mathbb I}}
\def\dbJ{{\mathbb J}}
\def\dbK{{\mathbb K}}
\def\dbL{{\mathbb L}}
\def\dbM{{\mathbb M}}
\def\dbN{{\mathbb N}}
\def\dbO{{\mathbb O}}
\def\dbP{{\mathbb P}}
\def\dbQ{{\mathbb Q}}
\def\dbR{{\mathbb R}}
\def\dbS{{\mathbb S}}
\def\dbT{{\mathbb T}}
\def\dbU{{\mathbb U}}
\def\dbV{{\mathbb V}}
\def\dbW{{\mathbb W}}
\def\dbX{{\mathbb X}}
\def\dbY{{\mathbb Y}}
\def\dbZ{{\mathbb Z}}

\def\Aut{{\rm Aut}}
\def\deg{{\rm deg}}

\def\la{{\langle}}
\def\ra{{\rangle}}

\bf\centerline{Homework Assignment \# 8. }\rm

{\bf Plan for the next week:} Galois correspondence (online lectures 20 and 21,
Section 14.2 in DF), maybe start finite fields (online lecture 22,
Section 14.3 in DF).
\skv
\bf\centerline{Problems, to be submitted by Thu, March 29th. }\rm
\skv
{\bf Problem 1:} \rm
Prove the interesting part of Corollary 18.7 form online notes: if $K/L/F$ is a tower of algebraic extensions and $K/L$ and $L/F$ are separable, then $K/F$ is separable (see online notes for a hint). Give a detailed argument.
\skv
\skv
{\bf Problem 2:} \rm  Let $p$ be a prime, $n\geq 2$ an integer, $f(x)=x^n-p$, and let
$K\subset \dbC$ be the splitting field for $f(x)$ over $\dbQ$.
Recall that $K=\dbQ(\sqrt[n]{p},\omega_n)$ where
$\omega_n=e^{2\pi i/n}$.
\begin{itemize}
\item[(a)] Describe the elements of $\Gal(K/\dbQ)$ by their actions on $\sqrt[n]{p}$
and $\omega_n$.
\item[(b)] Let $G=\dbZ/n\dbZ\rtimes (\dbZ/n\dbZ)^{\times}$ where 
$(\dbZ/n\dbZ)^{\times}$
acts on $\dbZ/n\dbZ$ by multiplication. Prove that $\Gal(K/\dbQ)$ is isomorphic
to a subgroup of $G$ and describe an explicit embedding. 
\item[(c)] (optional) Under which conditions on $n$ and $p$ is this embedding an isomorphism? 
You are not asked for necessary and sufficient conditions; just the most general sufficient
conditions that you can justify.
\end{itemize}

{\bf Problem 3:} \rm Let $F$ be a field, let $f(x)\in F[x]$
be an irreducible (over $F$) separable polynomial of degree $n$,
and let $K$ be a splitting field of $f(x)$.

\begin{itemize}
\item[(a)] Prove that $|\Gal(K/F)|$ is a multiple of $n$ and divides $n!$

\item[(b)] Let $n=3$. Prove that $\Gal(K/F)$ is isomorphic to either $\dbZ/3\dbZ$ or $S_3$.

\item[(c)] Let $n=4$ and assume that $|\Gal(K/F)|=8$.
Determine the isomorphism class of $\Gal(K/F)$.
\end{itemize}


{\bf Problem 4:} Let $f(x)\in \dbQ[x]$ be an irreducible polynomial
of degree $n$, and let $K$ be a splitting field of $f(x)$.
Label the roots of $f(x)$ by $1,\ldots, n$ (in some order),
and let $\iota: \Gal(K/\dbQ)\to S_n$ be the associated embedding.

\begin{itemize}
\item[(a)] Assume $f(x)$ has at least one non-real root. Prove
that the complex conjugation is an element of $\Gal(K/\dbQ)$
of order $2$.

\item[(b)] Assume that $f(x)$ has precisely two non-real roots. Prove
that the image of the complex conjugation under the embedding $\iota$
is a transposition.

\item[(c)] Suppose that $n=\deg(f)$ is prime and again assume that $f(x)$ has precisely
two non-real roots. Prove that $\Gal(K/\dbQ)$ is isomorphic to $S_n$.
{\bf Hint:} $\Gal(K/\dbQ)$ must contain an element of order $n$ (why?)

\end{itemize}
{\bf Problem 5:} The main goal of this problem is to give a slightly different solution
to Problem~4(c) in HW\#6. So, let $S=\{n_1,\ldots, n_k\}$ be a finite set of positive integers $\geq 2$ and $K=\dbQ(\sqrt{n_1},\ldots, \sqrt{n_k})$.

\begin{itemize}
\item[(a)] Prove that there exists a subset $T$ of $S$ such that the set
$P(T)=\{1\}\cup \{\sqrt{n} : n \mbox{ is a product of distinct elements of }T\}$
is a basis for $K/\dbQ$.

\item[(b)] Let $T$ be as in (a). Let $\alpha\in K$ be such that $\sigma(\alpha)=\pm \alpha$ for every $\sigma\in \Gal(K/\dbQ)$. Prove that $\alpha=q\beta$ for some $\beta\in T$ and $q\in\dbQ$.

\item[(c)] Assume now that $n_1,\ldots, n_k$ are pairwise relatively prime and none
of them is a square. Use (b) to give a short proof of the fact that $[K:\dbQ]=2^k$. 
\end{itemize}

\end{document} 